pdp: An R Package for Constructing Partial Dependence Plots

### 2017-01-01 manuscript/reviews/response.Rmd

### Response 1.4
This has been fixed in `pdp` (version 0.3.0). This was simply a side effect of `partial`'s internal use of `plyr::adply` which uses default variable names (e.g., `V1`, `V2`, etc.). This caused the column `V1` in the output data frame to be overwritten, producing the error. 
```{r, eval=TRUE}
library(pdp)
data <- data.frame(V1 = 1:10, V2 = 1:10)
mod <- lm(V2 ~ V1, data = data)
partial(mod, pred.var = "V1")  # no more error
```

### Comment 1.5
I think you should mention in the documentation something about the assumptions you make about accessing the model data. The following (somewhat contrived) example fails with
`"Error in train[[x]] : subscript out of bounds":`
```{r, eval=FALSE}
data(trees)
foo <- function(data)
{
    lm(Volume ~ ., data = data)
}
mod <- foo(trees)
library(pdp)
partial(mod, pred.var = "Girth", plot = TRUE)  # fails
```
and the following silently gives misleading results:

```{r, eval=FALSE}
library(pdp)
data(trees)
mod <- lm(Volume ~ ., data = trees)
partial(mod, pred.var = "Girth", plot = TRUE)  # ok
trees <- trees[1:2, ]
partial(mod, pred.var = "Girth", plot = TRUE)  # wrong xlim
```
I'm not necessarily saying that you need to resolve the above two
issues, but at least mention in the documentation that this kind of
thing can happen, and how to avoid it (basically call partial in the
same environment used to build the model, and don't change any of the
data used to build the model).

### Response 1.5
Agreed! The documentation for `partial` has also been expanded to discuss this. Unfortunately, some functions don't store a copy of the training data. However, I have expanded the code to work harder---especially for `S4` methods---to inform the user to supply the training data when necessary.

@@

### Comment 1.6
The following code gives an obscure error message:
```
"Error in seq.default(from = min(train[[x]], na.rm = TRUE),
   to = max(train[[x]],  :
  'from' cannot be NA, NaN or infinite"
```
Some better handholding for the user may be helpful here.
```{r, eval=FALSE}
library(pdp)
data(trees)
mod <- lm(Volume~., data=trees)
partial(mod, pred.var="nosuchvariable", plot=TRUE) # obscure err msg
```

### Response 1.6
This has been fixed in `pdp` (version 0.3.0) by adding an informative error message listing which variables are the cause of the issue.
```{r, eval=FALSE}
library(pdp)
data(trees)
mod <- lm(Volume~., data=trees)
partial(mod, pred.var=c("Girth", "foo", "bar"), plot=TRUE)
```
```
Error in partial.default(mod, pred.var = c("Girth", "foo", "bar"), plot = TRUE) : foo, bar not found in the training data.
```

### Comment 1.7
On page 2 of the paper it says "The columns of the data frame are labeled in the same order as the features supplied to `pred.var`, and the last column is always labeled `y`". What happens if `y` is one of the predictors?

### Response 1.7
The last column is now labeled `yhat`, which is less likely to be a predictor name. When one of the predictors is labeled `"y"` `randomForest::partialPlot` fails while `gbm::plot.gbm` does not. I think regardless of what the default column name is, there is always the possibility of this issue. This has been filed as an issue on the `pdp` GitHib page: https://github.com/bgreenwell/pdp/issues/22.

### 2017-01-01 manuscript/greenwell.tex
minor edits

\citet{harrison-1978-hedonic} were among the first to analyze the well-known Boston housing data. One of their goals was to find a housing value equation using data on median housing values from $n = 506$ census tracts in the suburbs of Boston from the 1970 census; see Table IV of \citet{harrison-1978-hedonic} for a description of each variable. The data violate many classical assumptions like linearity, normality, and constant variance. Nonetheless, \citeauthor{harrison-1978-hedonic}---using a combination of transformations, significance testing, and grid searches---were able to find a reasonable fitting model ($R^2 = 0.81$). Part of the payoff for there time and efforts was an interpretable prediction equation which is reproduced in Equation~\eqref{eqn:boston}.
\begin{equation}
\label{eqn:boston}
\begin{aligned}
\widehat{\log\left(MV\right)} &= 9.76 + 0.0063 RM^2 + 8.98\times10^{-5} AGE - 0.19\log\left(DIS\right) + 0.096\log\left(RAD\right) \\
  & \quad - 4.20\times10^{-4} TAX - 0.031 PTRATIO + 0.36\left(B - 0.63\right)^2 - 0.37\log\left(LSTAT\right) \\
  & \quad - 0.012 CRIM + 8.03\times10^{-5} ZN + 2.41\times10^{-4} INDUS + 0.088 CHAS \\
  & \quad - 0.0064 NOX^2
\end{aligned}
\end{equation}

@@

\begin{algorithm}
\begin{enumerate}
  \item For $i \in \left\{1, 2, \dots, k\right\}$:
  \begin{enumerate}
    \item Copy the training data and replace the original values of $x_1$ with the constant $x_{1i}$.
    \item Compute the vector of predicted values from the modified copy of the training data.
    \item Compute the average prediction to obtain $\bar{f}_1\left(x_{1i}\right)$.
  \end{enumerate}
  \item Plot the pairs $\left\{x_{1i}, \bar{f}_1\left(x_{1i}\right)\right\}$ for $i = 1, 2, \dotsc, k$.
\end{enumerate}
\caption{A simple algorithm for consructing the partial dependence of the response on a single predictor $x_1$. \label{alg:pdp}}
\end{algorithm}
Algorithm~\ref{alg:pdp} can be quite computationally intensive since it involves $k$ passes over the training records. Fortunately, this can be parallelized quite easily (more on this in Section~\ref{sec:computational}). This algorithm can also be easily extended to larger subsets of two or more features as well.

Limited implementations of Friedman's PDPs are available in packages \CRANpkg{randomForest} \citep{randomForest-pkg} and \CRANpkg{gbm}, among others; these are all limited in the sense that they only apply to the models fit using the respective package. For example, the \code{partialPlot} function in \pkg{randomForest} only applies to objects of class \code{"randomForest"} and the \code{plot} function in \pkg{gbm} only applies to \code{"gbm"} objects. While the \pkg{randomForest} implementation will only allow for a single predictor, the \pkg{gbm} implementation can deal with any subset of the predictor space. Partial dependence functions are not restricted to tree-based models; they can be applied to any supervised learning algorithm (e.g., multiple linear regression and neural networks). However, to our knowledge, there is no general package for constructing PDPs  in R. For example, PDPs for a \dfn{conditional random forest} as implemented by the \code{cforest} function in the \CRANpkg{party} and \CRANpkg{partykit} packages; see \citet{party-pkg} and \citet{partykit-pkg}, respectively. The \CRANpkg{pdp} \citep{pdp-pkg} package tries to close this gap by offering a general framework for constructing PDPs that can be applied to several classes of fitted models.

The \CRANpkg{plotmo} package \citep{plotmo-pkg} is one alternative to \pkg{pdp}. According to its author, \pkg{plotmo} constructs "a poor man's partial dependence plot." In particular, it plots a model's response when varying one or two predictors while holding the other predictors in the model constant (continuous features are fixed at their median value, while factors are held at their first level). These plots allow for up to two variables at a time. They are also less accurate than PDPs, but are faster to construct. For additive models (i.e., models with no interactions), these plots are identical in shape to PDPs.

@@

Many other techniques exist for visualizing relationships between the predictors and the response based on a fitted model. For example, the \CRANpkg{car} package \citep{fox-car-2011} contains many functions for constructing partial-residual and marginal-model plots. The \CRANpkg{effects} package \citep{fox-effects-2003} is also of interest. It provides tabular and graphical displays for the terms in parametric models. However, these methods were designed for simpler parametric models (e.g., linear and generalized linear models), whereas \pkg{plotmo}, \pkg{ICEbox}, and \pkg{pdp} are more useful for black box models (although, they can be used for simple parametric models as well).

@@

The \code{partial} function evaluates the partial dependence \eqref{eqn:pdf} from a fitted model over a grid of predictor values; the fitted model and predictors are specified using the \code{object} and \code{pred.var} arguments, respectively---these are the only required arguments. If \code{plot = FALSE} (the default), \code{partial} returns an object of class \code{"partial"} which inherits from the class \code{"data.frame"}; put another way, by default, \code{partial} returns a data frame with an additional class that is recognized by the \code{plotPartial} function. The columns of the data frame are labeled in the same order as the features supplied to \code{pred.var}, and the last column is always labeled \code{yhat}\footnote{There is one exception to this. When a function supplied via the \code{pred.fun} argument (this argument is discussed in Section~\ref{sec:prediction}) returns multiple predictions, the last two columns will be labeled \code{yhat} and \code{yhat.id}.} and contains the values of the partial dependence function $\bar{f}_s\left(\boldsymbol{z}_s\right)$. If \code{plot = TRUE}, then \code{partial} makes an internal call to \code{plotPartial} and returns the PDP in the form of a \CRANpkg{lattice} plot \citep{lattice-pkg}.

The \code{plotPartial} function can be used for displaying more advanced PDPs; it operates on objects of class \code{"partial"} and has many useful plotting options. For example, \code{plotPartial} makes it straight forward to add a LOESS smooth, or produce a 3-D surface instead of a false color level plot (the default). Of course, since the default output produced by \code{partial} is still a data frame, the user can easily use any plotting package he/she desires to visualize the results---\CRANpkg{ggplot2} \citep{ggplot2-pkg}, for instance.

\strong{Note:} as mentioned above, \pkg{pdp} relies on \pkg{lattice} for its graphics. \pkg{lattice} itself is built on top of \CRANpkg{grid} \citep{grid-pkg}. \pkg{grid} graphics behave a little differently than traditional R graphics, and two points are worth making (see \code{?lattice} for more details):
\begin{enumerate}
  \item \pkg{lattice} functions return a \code{"trellis"} object, but do not display it; the \code{print} method produces the actual display. However, due to R's automatic printing rule, the result is automatically printed when using these functions in the command line. If \code{plotPartial} is called inside of \code{source} or inside a loop (e.g., \code{for} or \code{while}), an explicit \code{print} statement is required to display the resulting graph; hence, the same is true when using \code{partial} with \code{plot = TRUE}.
  \item Setting graphical parameters via \code{par} typically has no effect on \pkg{lattice} plots. Instead, \pkg{lattice} provides its own \code{trellis.par.set} function for modifying graphical parameters.
\end{enumerate}
A consequence of the second point is that the \code{par} function cannot be used to control the layout of multiple \pkg{lattice} (and hence \pkg{pdp}) plots. Simple solutions are available in packages \CRANpkg{latticeExtra} \citep{latticeExtra-pkg} and \CRANpkg{gridExtra} \citep{gridExtra-pkg}. For convenience, \pkg{pdp} imports the \code{grid.arrange} function from \pkg{gridExtra} which makes it easy to display multiple \pkg{pdp} plots on a single page.

Currently supported models are described in Table~\ref{tab:models}. In a lot of these cases, \code{partial} should be able to automatically determine appropriate values for the options \code{type} (i.e., \code{"regression"} or \code{"classification"}) and \code{train} (the original training data). In other situations, the user may need to specify one or both of these additional arguments. These two arguments allow \code{partial} to be flexible enough to handle many of the model types not listed in Table~\ref{tab:models}; for example, neural networks from the \CRANpkg{nnet} package \citep{venables-modern-2002} and projection pursuit regression \citep{friedman-ppr-1981} using the \code{ppr} function in the \pkg{stats} package.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Multi-predictor PDPs
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Multi-predictor PDPs}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Avoiding extrapolation
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Avoiding extrapolation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Classification problems
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Classification problems}
\label{sec:class}

@@

Using the \code{pred.fun} argument, it is possible to obtain PDPs for classification problems on the probability scale. We just need to write a function that computes the predicted class probability of interest averaged across all observations. The function below can be used with the fitted SVM from the iris example of Section~\ref{sec:class} to extract the average predicted probability of belonging to the \code{Setosa} class.
\begin{example}
pred.prob <- function(object, newdata) {  # see ?predict.svm
  pred <- predict(object, newdata, probability = TRUE)
  prob.setosa <- attr(pred, which = "probabilities")[, "setosa"]
  mean(prob.setosa)
}
\end{example}
Next, we simply pass this function via the \code{pred.fun} argument in the call to \code{partial}. The following chunk of code construct PDPs for \code{Petal.Width} and \code{Petal.Length} on the probability scale. The results are displayed in Figure~\ref{fig:partial_iris_prob}.
\begin{example}
# PDPs for Petal.Width and Petal.Length on the probability scale
pdp.pw <- partial(iris.svm, pred.var = "Petal.Width", pred.fun = pred.prob,
                  plot = TRUE)
pdp.pl <- partial(iris.svm, pred.var = "Petal.Length", pred.fun = pred.prob,
                  plot = TRUE)
pdp.pw.pl <- partial(iris.svm, pred.var = c("Petal.Width", "Petal.Length"),
                     pred.fun = pred.prob, plot = TRUE)

>>>>

\citet{harrison-1978-hedonic} were among the first to analyze the well-known Boston housing data. One of their goals was to find a housing value equation using data on median housing values from $n = 506$ census tracts in the suburbs of Boston from the 1970 census; see \citet[Table IV]{harrison-1978-hedonic} for a description of each variable. The data violate many classical assumptions like linearity, normality, and constant variance. Nonetheless, \citeauthor{harrison-1978-hedonic}---using a combination of transformations, significance testing, and grid searches---were able to find a reasonable fitting model ($R^2 = 0.81$). Part of the payoff for there time and efforts was an interpretable prediction equation which is reproduced in Equation~\eqref{eqn:boston}.
\begin{equation}
\label{eqn:boston}
\begin{aligned}
\widehat{\log\left(MV\right)} &= 9.76 + 0.0063 RM^2 + 8.98\times10^{-5} AGE - 0.19\log\left(DIS\right) + 0.096\log\left(RAD\right) \\
  & \quad - 4.20\times10^{-4} TAX - 0.031 PTRATIO + 0.36\left(B - 0.63\right)^2 - 0.37\log\left(LSTAT\right) \\
  & \quad - 0.012 CRIM + 8.03\times10^{-5} ZN + 2.41\times10^{-4} INDUS + 0.088 CHAS \\
  & \quad - 0.0064 NOX^2
\end{aligned}
\end{equation}

@@

\begin{algorithm}
\begin{enumerate}
  \item For $i \in \left\{1, 2, \dots, k\right\}$:
  \begin{enumerate}
    \item Copy the training data and replace the original values of $x_1$ with the constant $x_{1i}$.
    \item Compute the vector of predicted values from the modified copy of the training data.
    \item Compute the average prediction to obtain $\bar{f}_1\left(x_{1i}\right)$.
  \end{enumerate}
  \item Plot the pairs $\left\{x_{1i}, \bar{f}_1\left(x_{1i}\right)\right\}$ for $i = 1, 2, \dotsc, k$.
\end{enumerate}
\caption{A simple algorithm for consructing the partial dependence of the response on a single predictor $x_1$. \label{alg:pdp}}
\end{algorithm}
Algorithm~\ref{alg:pdp} can be quite computationally intensive since it involves $k$ passes over the training records. Fortunately, the algorithm can be parallelized quite easily (more on this in Section~\ref{sec:computational}). It can also be easily extended to larger subsets of two or more features as well.

Limited implementations of Friedman's PDPs are available in packages \CRANpkg{randomForest} \citep{randomForest-pkg} and \CRANpkg{gbm} \citep{gbm-pkg}, among others; these are limited in the sense that they only apply to the models fit using the respective package. For example, the \code{partialPlot} function in \pkg{randomForest} only applies to objects of class \code{"randomForest"} and the \code{plot} function in \pkg{gbm} only applies to \code{"gbm"} objects. While the \pkg{randomForest} implementation will only allow for a single predictor, the \pkg{gbm} implementation can deal with any subset of the predictor space. Partial dependence functions are not restricted to tree-based models; they can be applied to any supervised learning algorithm (e.g., generalized additive models and neural networks). However, to our knowledge, there is no general package for constructing PDPs in R. For example, PDPs for a conditional random forest as implemented by the \code{cforest} function in the \CRANpkg{party} and \CRANpkg{partykit} packages; see \citet{party-pkg} and \citet{partykit-pkg}, respectively. The \CRANpkg{pdp} \citep{pdp-pkg} package tries to close this gap by offering a general framework for constructing PDPs that can be applied to several classes of fitted models.

The \CRANpkg{plotmo} package \citep{plotmo-pkg} is one alternative to \pkg{pdp}. According to \citeauthor{plotmo-pkg}, \pkg{plotmo} constructs "a poor man's partial dependence plot." In particular, it plots a model's response when varying one or two predictors while holding the other predictors in the model constant (continuous features are fixed at their median value, while factors are held at their first level). These plots allow for up to two variables at a time. They are also less accurate than PDPs, but are faster to construct. For additive models (i.e., models with no interactions), these plots are identical in shape to PDPs. As of \pkg{plotmo} version 3.3.0, there is now support for constructing PDPs, but it is not the default. The main difference is that \pkg{plotmo}, rather than applying step 1. (a)-(c) in Algorithm~\ref{alg:pdp}, accumulates all the data at once thereby reducing the number of internal calls to \code{predict}. The trade-off is a slight increase in speed at the expense of using more memory. So, why use the \pkg{pdp} package? As will be discussed in the upcoming sections, \pkg{pdp}:
\begin{itemize}
  \item contains only a few functions with relatively few arguments;
  \item does \underline{not} produce a plot by default;
  \item can be used more efficiently with \code{"gbm"} objects (see Section~\ref{sec:computational});
  \item produces graphics based on \CRANpkg{lattice} \citep{lattice-pkg}, which are far more flexible than base R graphcs;
  \item defaults to using false color level plots for multivariate displays (see Section~\ref{sec:multi});
  \item contains options to mitigate the risks associated with extrapolation (see Section~\ref{sec:computational});
  \item has the option to construct PDPs in parallel (see Section~\ref{sec:computational});
    \item is extremely flexible in the types of PDPs that can be produced (see Section~\ref{sec:prediction}),
\end{itemize}

@@

Many other techniques exist for visualizing relationships between the predictors and the response based on a fitted model. For example, the \CRANpkg{car} package \citep{fox-car-2011} contains many functions for constructing \dfn{partial-residual} and \dfn{marginal-model} plots. \dfn{Effect displays}, available in the \CRANpkg{effects} package \citep{fox-effects-2003}, provides tabular and graphical displays for the terms in parametric models while holding all other predcitors at some constant value---similar in spirit to \pkg{plotmo}'s marginal model plots. However, these methods were designed for simpler parametric models (e.g., linear and generalized linear models), whereas \pkg{plotmo}, \pkg{ICEbox}, and \pkg{pdp} are more useful for black box models (although, they can be used for simple parametric models as well).

@@

The \code{partial} function evaluates the partial dependence \eqref{eqn:pdf} from a fitted model over a grid of predictor values; the fitted model and predictors are specified using the \code{object} and \code{pred.var} arguments, respectively---these are the only required arguments. If \code{plot = FALSE} (the default), \code{partial} returns an object of class \code{"partial"} which inherits from the class \code{"data.frame"}; put another way, by default, \code{partial} returns a data frame with an additional class that is recognized by the \code{plotPartial} function. The columns of the data frame are labeled in the same order as the features supplied to \code{pred.var}, and the last column is labeled \code{yhat}\footnote{There is one exception to this. When a function supplied via the \code{pred.fun} argument returns multiple predictions, the second to last and last columns will be labeled \code{yhat} and \code{yhat.id}, respectively (see Section~\ref{sec:prediction}).} and contains the values of the partial dependence function $\bar{f}_s\left(\boldsymbol{z}_s\right)$. If \code{plot = TRUE}, then \code{partial} makes an internal call to \code{plotPartial} (with fewer plotting options) and returns the PDP in the form of a \pkg{lattice} plot (i.e., a \code{"trellis"} object). \strong{Note:}  it is reccomended to call \code{partial} with \code{plot = FALSE} and store the results; this allows for more flexible plotting, and the user will not have to waste time calling \code{partial} again if the default plot is not sufficient.

The \code{plotPartial} function can be used for displaying more advanced PDPs; it operates on objects of class \code{"partial"} and has many useful plotting options. For example, \code{plotPartial} makes it straight forward to add a LOESS smooth, or produce a 3-D surface instead of a false color level plot (the default). Of course, since the default output produced by \code{partial} is still a data frame, the user can easily use any plotting package he/she desires to visualize the results---\CRANpkg{ggplot2} \citep{ggplot2-pkg}, for instance (see Section~\ref{sec:classification} and Section~\ref{sec:prediction} for examples).

\strong{Note:} as mentioned above, \pkg{pdp} relies on \pkg{lattice} for its graphics. \pkg{lattice} itself is built on top of \CRANpkg{grid} \citep{grid-pkg}. \pkg{grid} graphics behave a little differently than traditional R graphics, and two points are worth making (see \code{?lattice} for more details):
\begin{enumerate}
  \item \pkg{lattice} functions return a \code{"trellis"} object, but do not display it; the \code{print} method produces the actual display. However, due to R's automatic printing rule, the result is automatically printed when using these functions in the command line. If \code{plotPartial} is called inside of \code{source} or inside a loop (e.g., \code{for} or \code{while}), an explicit \code{print} statement is required to display the resulting graph; hence, the same is true when using \code{partial} with \code{plot = TRUE}.
  \item Setting graphical parameters via \code{par} typically has no effect on \pkg{lattice} plots. Instead, \pkg{lattice} provides its own \code{trellis.par.set} function for modifying graphical parameters.
\end{enumerate}
A consequence of the second point is that the \code{par} function cannot be used to control the layout of multiple \pkg{lattice} (and hence \pkg{pdp}) plots. Simple solutions are available in packages \CRANpkg{latticeExtra} \citep{latticeExtra-pkg} and \CRANpkg{gridExtra} \citep{gridExtra-pkg}. For convenience, \pkg{pdp} imports the \code{grid.arrange} function from \pkg{gridExtra} which makes it easy to display multiple \pkg{grid}-based graphical objects on a single plot (these include graphics produced using \pkg{lattice} (hence, \pkg{pdp}) and \pkg{ggplot2}). This is demonstrated in multiple examples throughout this paper.

Currently supported models are described in Table~\ref{tab:models}. In a lot of these cases, \code{partial} should be able to automatically determine appropriate values for the options \code{type} (i.e., \code{"regression"} or \code{"classification"}) and \code{train}\footnote{For reasons described in the package documentation, it is good practice to always supply the training data via the \code{train} argument in the call to \code{partial}; see \code{?partial for details}} (the original training data). In other situations, the user may need to specify one or both of these additional arguments. These two arguments allow \code{partial} to be flexible enough to handle many of the model types not listed in Table~\ref{tab:models}; for example, neural networks from the \CRANpkg{nnet} package \citep{venables-modern-2002} and projection pursuit regression \citep{friedman-ppr-1981} using the \code{ppr} function in the \pkg{stats} package.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Multi-predictor PDPs
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Multi-predictor PDPs}
\label{sec:multi}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Avoiding extrapolation
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Avoiding extrapolation}
\label{sec:extrapolation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Classification problems
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Classification problems}
\label{sec:classification}

@@

Using the \code{pred.fun} argument, it is possible to obtain PDPs for classification problems on the probability scale. We just need to write a function that computes the predicted class probability of interest averaged across all observations. The function below can be used with the fitted SVM from the iris example of Section~\ref{sec:classification} to extract the average predicted probability of belonging to the \code{Setosa} class.
\begin{example}
pred.prob <- function(object, newdata) {  # see ?predict.svm
  pred <- predict(object, newdata, probability = TRUE)
  prob.setosa <- attr(pred, which = "probabilities")[, "setosa"]
  mean(prob.setosa)
}
\end{example}
Next, we simply pass this function via the \code{pred.fun} argument in the call to \code{partial}. The following chunk of code construct PDPs for \code{Petal.Width} and \code{Petal.Length} on the probability scale. The results are displayed in Figure~\ref{fig:partial_iris_prob}.
\begin{example}
# PDPs for Petal.Width and Petal.Length on the probability scale
pdp.pw <- partial(iris.svm, pred.var = "Petal.Width", pred.fun = pred.prob,
                  plot = TRUE)
pdp.pl <- partial(iris.svm, pred.var = "Petal.Length", pred.fun = pred.prob,
                  plot = TRUE)
pdp.pw.pl <- partial(iris.svm, pred.var = c("Petal.Width", "Petal.Length"),
                     pred.fun = pred.prob, plot = TRUE)
